---
title: "National Women’s Soccer League"
author: "YOUR NAME HERE"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
editor: visual
editor_options:
  chunk_output_type: console
---

# Getting Started (Data Wrangling)

## Loading Packages and Data

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(viridis)
library(kableExtra)
library(ggridges)
library(scales)
#install.packages("devtools")
devtools::install_github("nwslR/nwslR", ref = "v0.0.0.9002")
library(nwslR)
```

```{r}
#| label: load-data
#| message: false
matches <- load_matches()
players <- load_players()
teams <- load_teams()
metrics <- load_metrics()
```

```{r}
View(VAR)
#replace VAR with `players`, `teams`, and `metrics` to see the other data sets
```

## Exercise 1

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

## Loading Portland Thorns FC Matches 2023

```{r}
portland_thorns_id <- teams %>%
  filter(team_name == ID_VARIABLE_NAME) %>%
  pull(external_team_id)

#Replace ID_VARIABLE_NAME with the answer for Exercise 1.a
```

```{r}

portland_thorns_matches <- matches %>%
  filter((home_team_id == portland_thorns_id | away_team_id == portland_thorns_id) & season == "2023") %>%
  pull(match_id)

#Replace the MATCH_ID_NAME with the id of the matches where portland thorns fc played
#You got it right if in your environment you see 'chr [1:23] "portland-thorns-fc..."'

```

```{r}
# Load player stats for all 2023 Portland Thorns matches
all_thorns_player_stats <- map_dfr(
  portland_thorns_matches, 
  ~ load_player_match_stats(match_id = .x)  
) %>%
  filter(team_id == portland_thorns_id)

# Load team stats for all 2023 Portland Thorns matches
team_match_stats <- map_dfr(
  portland_thorns_matches, 
  ~ load_team_match_stats(match_id = .x)  
) %>%
  filter(team_id == portland_thorns_id & period == 'total')

#Load team season stats for Portland Thorns FC performance in 2023
portland_thorns_team_season_stats <- load_team_season_stats(
  team_id = "POR",
  season = "2023"
)
```

## Exercise 2

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

# Research Question 1

## Loading the defensive stats information from Portland Thorns FC

```{r}
fastbreak_stats <- all_thorns_player_stats %>%
  select(
    player_id, match_id, team_id, 
    contains("fastbreaks"),          
    contains("shots_fastbreak"), 
    contains("goals_fastbreak"), 
    VAR1, VAR2, VAR3     
  )

# Replace VAR1, VAR2, VAR3 with the variables that you think are important to analyze the effectiveness of a defensive lineup. You can also add a comma and add even more variables if you think they are important.

```

## Exercise 3

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

## Visualizing Portland Thorns FC fastbreak stats

```{r}
total_fastbreaks <- sum(DATA_SET_VAR$fastbreaks, na.rm = TRUE)
total_fastbreaks_shots <- sum(DATA_SET_VAR$shots_fastbreak, na.rm = TRUE)
total_fastbreaks_goals <- sum(DATA_SET_VAR$goals_fastbreak, na.rm = TRUE)

#Replace DATA_SET_VAR with the dataset table where we are getting the information from
```

```{r}
average_possession_pct <- mean(DATA_SET_VAR$possession_pct, na.rm = TRUE)
#Replace DATA_SET_VAR with the dataset table where we are getting the information from
```

```{r}
attempts_to_shots <- data.frame(
  stage = c("Fastbreak Attempts", "Shot Opportunities"),
  count = c(total_fastbreaks, total_fastbreaks_shots),
  conversion_rate = c(100, (total_fastbreaks_shots/total_fastbreaks)*100)
)

# Create visualization
ggplot(attempts_to_shots, aes(x = stage, y = count, fill = stage)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(count, "\n(", round(conversion_rate, 1), "%)")), 
            vjust = -0.5, size = 5) +
  scale_fill_manual(values = c("#8C489F", "#FFD700")) +
  labs(
    title = "Portland Thorns FC: Fastbreak to Shot Conversion (2023)",
    subtitle = paste("Conversion Rate:", round((total_fastbreaks_shots/total_fastbreaks)*100, 1), "%"),
    x = NULL,
    y = "Count",
    caption = "Data: NWSL (2023)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.x = element_text(size = 12),
    legend.position = "none"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
```

```{r}
shot_goal_data <- data.frame(
  stage = c("Shot Opportunities", "Goals"),
  count = c(total_fastbreaks_shots, total_fastbreaks_goals),
  conversion_rate = c(
    100,  # Baseline (all shots)
    (total_fastbreaks_goals / total_fastbreaks_shots) * 100  # Goals per shot
  )
)

# Plot
ggplot(shot_goal_data, aes(x = stage, y = count, fill = stage)) +
  geom_col(width = 0.6, fill = c("#FFD700", "#006341")) +
  geom_text(
    aes(label = paste0(count, "\n(", round(conversion_rate, 1), "%)")),
    vjust = -0.5, size = 5, color = "black"
  ) +
  labs(
    title = "Portland Thorns FC: Fastbreak Finishing (2023)",
    subtitle = "How many shot opportunities resulted in goals?",
    x = NULL,
    y = "Count",
    caption = "Data: NWSL (2023)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.x = element_text(size = 12),
    legend.position = "none"
  )
```

## Exercise 4

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

# Research Question 2

## Loading the attack stats information from Portland Thorns FC

```{r}
offensive_team_stats <- all_thorns_player_stats %>%
  select(
    player_id, match_id, team_id, 
    contains("big_chances_created"),    
    contains("big_chances_scored"), 
    contains("big_chances_missed"), 
    contains("contests_total"),
    contains("contests_won"),
    contains('pass_pct_opposition_half'),
    VAR1, VAR2, VAR3
  )

# Replace VAR1, VAR2, VAR3 with the variables that you think are important to analyze the effectiveness of a defensive lineup. You can also add a comma and add even more variables if you think they are important.

pass_pct <- portland_thorns_team_season_stats %>%
  pull(contains('pass_pct_opposition_half')) %>%
  as.numeric()

```

## Exercise 5

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

## Visualizing Portland Thorns FC offensive stats

```{r}
total_contests <- sum(DATA_SET_VAR$contests_total, na.rm = TRUE)
total_contests_won <- sum(DATA_SET_VAR$contests_won, na.rm = TRUE)
total_big_chances_created <- sum(DATA_SET_VAR$big_chances_created, na.rm = TRUE)
total_big_chances_scored <- sum(DATA_SET_VAR$big_chances_scored, na.rm = TRUE)
total_big_chances_missed <- sum(DATA_SET_VAR$big_chances_missed, na.rm = TRUE)

#Replace DATA_SET_VAR with the dataset table where we are getting the information from
```

```{r}
pass_pct_opp_half <- portland_thorns_team_season_stats %>%
  pull(contains('pass_pct_opposition_half')) %>%
  as.numeric()
```

```{r}
contest_data <- data.frame(
  category = c("Contests Won", "Contests Lost"),
  count = c(total_contests_won, total_contests - total_contests_won)
)

ggplot(contest_data, aes(x = "Contests", y = count, fill = category)) +
  geom_col() +
  geom_text(aes(label = count), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("#006341", "#8C489F")) +  # Thorns colors
  labs(
    title = "Portland Thorns FC: Contest Success (2023)",
    subtitle = paste0("Win rate: ", round(total_contests_won / total_contests * 100, 1), "%"),
    x = NULL,
    y = "Total Contests",
    fill = "Outcome"
  ) +
  theme_minimal()
```

```{r}
big_chance_data <- data.frame(
  stage = c("Created", "Scored"),
  count = c(total_big_chances_created, total_big_chances_scored)
)

ggplot(big_chance_data, aes(x = stage, y = count, fill = stage)) +
  geom_col() +
  geom_text(aes(label = abs(count)), vjust = -0.5) +
  scale_fill_manual(values = c("#FFD700", "#8C489F")) +
  labs(
    title = "Big Chances: Creation vs. Conversion (2023)",
    subtitle = paste0("Conversion rate: ", round(total_big_chances_scored / total_big_chances_created * 100, 1), "%"),
    x = NULL,
    y = "Count"
  ) +
  theme_minimal()
```

## Exercise 6

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

# Wrap-up / reflection

## Exercise 7

a.  YOUR ANSWER HERE

b.  YOUR ANSWER HERE

c.  YOUR ANSWER HERE

d.  YOUR ANSWER HERE

# BONUS

## Exercise 8

Bonus. YOUR ANSWER HERE
